---
title: "covid 19"
author: "Rachel Polack"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

Examining the table, it seems that more men have died in this study as compared to females (44 and 25 respectively). Additionally, there seems to be more females in the study overall (1831) compared to males (1411); this would suggest that a higher proportion of men died compared to women in this study, but further testing is required to confirm whether this difference in proportion is statistically significant. 

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

When considering the frequency-based ggplot bar charts computed, it appears that there are about 1800 living females in the study, and about 1400 living males. However more male persons were reported as dead than females, even though there were hundreds more females as a whole in the study. Regarding the proportion based bar chart, a majority of the reported deaths were indicated as male. Further studying these proportions requires a numerical analysis.
`

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The first table table created examines the data in terms of totals of results. It seems that a solid majority of those reported "dead" as a result were males (64%). In analyzing the second table, which was created on the basis of total of gender, a higher proportion of men (3%) were reported dead, as compared to only 1.4% of females. While inferential results are required to assess the statistical significance of these results, it seems that there may be some type of relationship between death and sex in this study, when entertaining the data accumulated thus far. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Running the chi-squared test confirms the frequency that we had estimated by running our graphical and numerical tests (see "observed counts"). However, the clearest explanation of our results comes in the form of our fisher exact test, which generated a p-value of .0009 for the data provided. Furthermore, the odds ratio (2.3) can be interpreted as the notion that the males are 2.3 times more likely to die from Covid, according to the results of this study.  

## Conclusions

In summary, this assignment ventured to explain whether or not death by Covid was independent of sex, in a study of individuals from South Korea. We sought to explain from graphical, numerical, and inferential stances. Our proportion-based bar charts shows a variation in proportion among those who died (favoring men), and this furthered by our numerical results which provided the exact percentage to fortify that claim. Finally, our inferential results provided the exact counts of the prior tests, but also yielded a low p-value, helping to certify that our results had not purely occurred through by chance. Our odds ratio (2.3) was the single most compelling piece of data generated, helping to reinforce the alternative hypothesis that dying is actually dependent (at least in part) on sex.  

