---
title: "covid 19"
author: "Rachel Polack"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

Examining the table, it seems that more men have died in this study as compared to females (44 and 25 respectively). Additionally, there seems to be more females in the study overall (1831) compared to males (1411); this would suggest that a higher proportion of men died compared to women in this study, but further testing is required to confirm whether this difference in proportion is statistically significant. 

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`

  





  
## Numerical results


## Inferential results 




## Conclusions

